---
title: "Practical Machine Learning"
author: "Abanoub Amgad"
date: "December 26, 2015"
output: html_document
---
In This Project, i'll try to build a model for the PML data to predict activity quality from activity monitors  

###Reading & Cleaning Data

```{r}
library(caret)
library(randomForest)
set.seed(12345)
pml <- read.csv("c:/Users/abano/Downloads/pml-training.csv",stringsAsFactors = F)
```
after reading the data and running str(pml), i found out that some variables are simply not available, some are empty and some won't be really useful in prediction (like the names)  
so, to create a better model i had to remove them.  
```{r}
inTrain <- createDataPartition(as.factor(pml$classe),p=0.75,list=FALSE)
training <- pml[inTrain,]
validation <- pml[-inTrain,] #creating a validation set
cleanTraining <- training[,colSums(is.na(training))==0] #removing NAs
cleanTraining$classe <- as.factor(cleanTraining$classe) 
numericCols <- numeric()
for(i in 1:ncol(cleanTraining)){ #locating the empty variables
  if(class(cleanTraining[,i]) != "character"){
    numericCols <- c(numericCols,i)
  }
}
cleanTraining <- cleanTraining[,numericCols] #removing empty variables
cleanTraining <- cleanTraining[,-1:-4] #removing name, time as they won't affect classification
```

we are now ready to build our model, i'll choose random forests as it's a very good classification technique and it suits our case

###Using Random Forests

```{r}
fit <- randomForest(data=cleanTraining,classe~.,ntree = 50,proximity=T,importance=T)
```

after we got the model, we should calculate the accuracy & in sample error

###Comparing Errors

```{r}
inpred <- predict(fit,training)
confusionMatrix(table(inpred,as.factor(training$classe)))
```

an accuracy of 1 !! that was expected as this is the training data, on the contrary .. the accuracy of the validation set prediction is expected to be lower .. which means HIGHER error 

```{r}
outpred <- predict(fit,validation)
confusionMatrix(table(outpred,as.factor(validation$classe)))
```
as we expected, the accuracy of the validation set prediction is lower (higher error) .. but actually it's not that bad .. we still have a pretty good prediction accuracy that will get us the 20 points of the programming assignment! 
```{r}
test <- read.csv("c:/Users/abano/Downloads/pml-testing.csv",stringsAsFactors = F)
predict(fit,test)
```